---
title: "Using the Unit Cost Model"
author: "Alex Hill"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{predicting costs}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 8,
  fig.height = 8
)
```

```{r setup, include = FALSE}
library(capturetb)
library(ggplot2)
library(dplyr)
```

This package was developed for two reasons: firstly, to make the model development process transparent and reproducible, and secondly, to allow people to make use of the final model to predict TB outpatient costs. This vignette speaks to the second of these aims. If you want to use the model to predict costs, this vignette is the only documentation you need to read. 

## Model structure

The unit cost model is a mixed effects model with country-specific intercepts and covariate effects that are assumed shared across countries. Log costs are assumed to vary normally:

$$
\log(\text{USD\_unitcost\_total}_i) \sim \mathcal{N}(\mu_i, \sigma^2)
$$

where $\mu_i$ is given by:

$$
\mu_i = \alpha_{c_i} + \sum_{j=1}^J{\beta_j \times x_{ji}}
$$

The country-specific intercepts $\alpha_{c_i}$ are assumed to vary normally 
according to:

$$
\alpha_{c} \sim \mathcal{N}(\mu_\alpha, \sigma_\alpha^2)
$$

Model instances are created via the `unitcost` function:

```{r model-creation}
model <- capturetb::unitcost()

# See model covariates
covariates <- model$covariates()
print(covariates)

# See priors used in the model
priors <- model$priors()
print(priors)
```

The prior distribution for each model parameter ($\sigma$, each $\beta_j$, $\mu_\alpha$, and $\sigma_\alpha$) can be visualised by calling `plot(priors, parameter = "name_of_param")`:

```{r beta-priors, fig.width=10, fig.height=6}
library(gridExtra)
plots <- list()
for(i in 1:3) {
  plots[[i]] <- plot(priors, parameter = paste0("beta[", i, "]"))
}
plots[[4]] <- plot(priors, parameter = "mu_alpha")
plots[[5]] <- plot(priors, parameter = "sigma_alpha")
plots[[6]] <- plot(priors, parameter = "sigma")
do.call(grid.arrange, c(plots, ncol = 3))
```

The posterior (fitted) distribution of each parameter can be generated by calling `model$plot_posteriors`:

```{r}
#| fig-cap: Fitted beta cefficients
model$plot_posteriors(pars = paste0("beta[", 1:length(covariates), "]")) + 
  scale_y_discrete(labels = covariates)
```

```{r}
#| fig-cap: Fitted mu_alpha, sigma_alpha and sigma parameters
model$plot_posteriors(pars = c("mu_alpha", "sigma_alpha", "sigma",  paste0("alpha[", 1:5, "]")))
```

## Training data

The unit cost model is fitted using top-down economic costs from the ValueTB database. The raw training data are installed with the package and include costs for multiple different output types; run `capturetb::outputs()` to retrieve a complete list of available outputs:

```{r data-exploration}
available_outputs <- capturetb::outputs()
head(available_outputs)
```

Subsets of the data can be retrieved using `capturetb::get_data()`.

```{r load-data}
data <- capturetb::get_data("OP treatment visit")
summarised <- data |> 
  group_by(fc_country) |>
  summarise(n_obs = n())

knitr::kable(summarised)
```

The fitted model predicts the cost of one outpatient treatment visit, given key cost drivers. In the dataset this output is coded as "OP treatment visit". In the raw data some facilities have multiple entries; one of each of these was chosen at random 
to include in the training data to avoid biasing the estimates with highly correlated errors.

The final training data can be seen by calling `model$training_data`:

```{r}
dat <- model$training_data()
head(dat)

knitr::kable(dat |> 
  group_by(fc_country) |>
  summarise(n_obs = n())
)
```

## Model accuracy

To see how the model performs on the training data:

```{r predictions}
# Various measures of fit
performance <- model$performance(scale = "log")
knitr::kable(performance)
```

and by country: 

```{r country-analysis}
# Performance by country
country_performance <- model$performance(by_country = TRUE)
colnames(country_performance) <- c("Country", "MAE",
 "RMSE", "Correlation", "95% CI Coverage")

knitr::kable(country_performance)
```

Visusaling predictions for inputs in the training data, including credible intervals, against observed costs:

```{r prediction-intervals}
# Plot with prediction intervals
model$plot_fit(include_ci = TRUE, scale = "log")
```

## Generating predictions

Predictions of unit cost for new inputs, not in the training set, are generated via the `predict` method:

```{r}
# Include all covariates and facility country `fc_country`
new_inputs <- list(
  log_USD_p_bldgspace = 1,
  logVisits = 6.9, 
  logVisitsPP = -1.29, 
  secondary = FALSE, 
  urban = FALSE, 
  public = TRUE,
  fc_country = "Ethiopia"
)

pred <- model$predict(new_inputs, scale = "natural", summarised = TRUE)

# Expected unit cost is mean prediction
expected_unit_cost <- pred$mean
print(expected_unit_cost)

knitr::kable(pred)
```

To see the full distribution of predicted costs:

```{r}
pred <- model$predict(new_inputs, scale = "natural", summarised = FALSE)
ggplot(data.frame(cost = pred), aes(x = cost)) + geom_density()
```
