---
title: "Using the Unit Cost Model"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using the Unit Cost Model}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 8,
  fig.height = 8
)
```

```{r setup, include = FALSE}
library(capturetb)
library(ggplot2)
library(dplyr)
```

This package was developed for two reasons: firstly, to make the model development process transparent and reproducible, and secondly, to allow people to make use of the final models to predict TB outpatient visit costs. This vignette speaks to the second of these aims. If you want to use the models to predict costs, this vignette is the only documentation you need to read. 

## Model structure

The unit cost model is a mixed effects model with country, facility and visit type effects and covariate effects that are assumed shared across countries, facilities and visit types. Log costs are assumed to vary normally:

$$
\log(\text{USD\_unitcost\_total}_i) \sim \mathrm{N}(\mu_i, \sigma^2)
$$

where $\mu_i$ is given by:

$$
\mu_i = \alpha + \sum_{j=1}^J{\beta_j x_{ji}} + \gamma_{c_i} +  \delta_{f_i} +  \zeta_{v_i}
$$

where $x_{ji}$ represents the value of the jth covariate, $c_i$ the country, $f_i$ the facility, and $v_i$ the visit type of observation $i$. The variables $\gamma_{c_i}$, $\delta_{f_i}$ and $\zeta_{v_i}$ account for systematic differences across countries, facilities and visit types, respectively, and are assumed to vary normally about zero:

$$
\gamma_{c} \sim \mathrm{N}(0, \sigma_c^2)
$$

$$
\delta_{f} \sim \mathrm{N}(0, \sigma_f^2)
$$

$$
\zeta_{o} \sim \mathrm{N}(0, \sigma_v^2)
$$

Model instances are created via the `unitcost` function:

```{r model-creation}
model <- capturetb::unitcost()

# See model covariates
covariates <- model$covariates()
print(covariates)

# See priors used in the model
priors <- model$priors()
print(priors)
```

The prior distribution for each model parameter can be visualised by calling `plot(priors, parameter = "name_of_param")`:

```{r beta-priors, fig.width=10, fig.height=6}
library(gridExtra)
plots <- list()
for(i in 1:9) {
  plots[[i]] <- plot(priors, par = paste0("beta[", i, "]"))
}
plots[[10]] <- plot(priors, par = "alpha")
plots[[11]] <- plot(priors, par = "sigma")
plots[[12]] <- plot(priors, par = "sigma_c")
plots[[13]] <- plot(priors, par = "sigma_f")
plots[[14]] <- plot(priors, par = "sigma_v")

do.call(grid.arrange, c(plots, ncol = 3))
```

The posterior (fitted) distribution of each parameter can be generated by calling `model$plot_posteriors`:

```{r, fig-beta}
#| fig-cap: Fitted beta cefficients
model$plot_posteriors(pars = paste0("beta[", 1:length(covariates), "]")) + 
  scale_y_discrete(labels = covariates)
```

```{r, fig-other}
#| fig-cap: Fitted intercept and standard deviations of random effects
model$plot_posteriors(pars = c("alpha", "sigma", "sigma_c", "sigma_f", "sigma_v"))
```

## Training data

The `unitcost` model is fitted using top-down economic costs in 2018 International Dollars from the ValueTB database. The raw data are installed with the package and include costs for multiple different output types. Output types are also grouped into output groups. Run `capturetb::output_groups` to see all groups, and `capturetb::outputs()` to retrieve a complete list of available outputs:

```{r data-exploration}
available_output_groups <- capturetb::output_groups()
head(available_output_groups)

available_outputs <- capturetb::outputs()
head(available_outputs)
```

Subsets of the data can be retrieved using `capturetb::get_data()`, filtering by output group or output type as needed:

```{r load-data}
data <- capturetb::get_data(output_group = "OP")
summarised <- data |> 
  group_by(fc_country) |>
  summarise(n_obs = n())

knitr::kable(summarised)
```

The fitted model predicts the cost of a single outpatient visit, given key cost drivers. In the dataset outpatient visits of different types have output group "OP".

The training data was processed from the raw data by adding one additional variable, `n_services`, which is the number of distinct TB outpatient services provided at each facility, and by centering all numeric covariates by their sample mean. The training data can be inspected by calling `model$training_data()`:

```{r training-data}
dat <- model$training_data() |> 
	select(model$covariates())
head(dat)
```

## Model accuracy

To see how the model performs on the training data:

```{r predictions}
# Various measures of fit
performance <- model$performance(scale = "natural")
colnames(performance) <- c("MAE",
 "RMSE", "95% CI Coverage", "Median CI width", "Bayesian R-squ")
knitr::kable(performance)
```

and by country: 

```{r country-analysis}
# Performance by country
country_performance <- model$performance(by_country = TRUE)
colnames(country_performance) <- c("Country", "MAE",
 "RMSE", "95% CI Coverage", "Median CI width", "Bayesian R-squ")

knitr::kable(country_performance)
```

By default, performance is reported after marginalising over facility effects. To see the performance of the full conditional model, including known facility effects:

```{r conditional}
# Various measures of fit
performance <- model$performance(scale = "natural", conditional = TRUE)
colnames(performance) <- c("MAE",
 "RMSE", "95% CI Coverage", "Median CI width", "Bayesian R-squ")
knitr::kable(performance)
```

Visualisiing predictions for inputs in the training data, including credible intervals, against observed costs:

```{r prediction-intervals}
# Plot with prediction intervals
do.call(grid.arrange, 
	c(model$plot_fit(include_ci = TRUE, scale = "log") + ggtitle("Marginal model"),
  	model$plot_fit(include_ci = TRUE, scale = "log", conditional = TRUE) + ggtitle("Conditional model")))
```

## Generating predictions

Predictions of unit cost for new inputs, not in the training set, are generated via the `predict` method. Note that numeric covariates must be centered; to convert raw values to centered values, use the `prepare_covariates` utility:

```{r, prepare}
# Include all covariates plus facility country `fc_country` and visit type `output`
new_inputs <- list(
  log_ID_p_bldgspace = 1, 
  logVisits = 6.9, 
  logVisitsPP_TB = -1.29, 
	n_services = 1,
	primary = FALSE,
  secondary = FALSE, 
	tertiary = FALSE,
  urban = FALSE, 
  public = TRUE,
  fc_country = "Ethiopia",
	output = "op_treatmentvisit"
)

# This will center covariates as required by the model
covariates <- capturetb::prepare_covariates(raw = new_inputs, mod = model)
```

```{r, pred}
pred <- model$predict(covariates,
 scale = "natural",
 summarised = TRUE,
 centrality = c("mean", "median"),
 ci = c(0.95, 0.9, 0.8))

# Expected unit cost is mean prediction
expected_unit_cost <- pred$Mean
print(expected_unit_cost)

knitr::kable(pred)
```

To see the full distribution of predicted costs:

```{r, plot-pred}
pred <- model$predict(covariates, scale = "natural", summarised = FALSE)
ggplot2::ggplot(data.frame(cost = pred), ggplot2::aes(x = cost)) + ggplot2::geom_density()
```
