---
title: "Model Fitting"
author: "Alex Hill"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Model Fitting}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

```{r setup, message = FALSE, warn = FALSE}
library(capturetb)
library(ggplot2)
library(dplyr)
```

This vignette demonstrates how the default model was fitted and sensitivity analyses performed. It also shows how models could be fit using different covariates, target variable, training data or priors.

The following functions were included in the package to make the process of model development transparent and reproducible, but if you want to use the final CaptureTB model to predict costs you should read `vignette("01_model-predictions")` instead.

## Creating a model instance

A model instance requires a list of covariates, a target variable to predict, training data, and priors for parameters. 

```{r model-inputs}
covariates <- c("log_USD_p_bldgspace",
 "logVisits", "logVisitsPP", "logVisitsPP_TB", "primary",
 "secondary", "urban", "public")

# Specifying priors for the fixed effects
# One beta coefficient for each covariate
# Other parameters will take default values
priors <- capturetb::capturetb_priors(
  beta.mean = c(0,-0.0142, -0.0412, -0.0412, 0.304, 0.348, 0.352, -0.29),
  beta.precision = c(0.01, 227, 139, 139, 46.4, 80.8, 55.2, 234.5)
)

data <- get_data("OP diagnostic visit") 
# or provide your own data;
# see capturetb::outputs() for all output data 
# installed with the package
```

We now create an instance of the `capturetb::MixedEffects` class to fit a model with country-specific intercepts and shared slopes. To fit a fixed effects model we could use the FixedEffects class, or to fit a model with random slopes, 
we could use a RandomSlopes class. See `vignette("03_model-comparisons")` for details on the different model structures.
 
```{r model-creation}
model <- capturetb::MixedEffects$new(
  dat = data,
  covariates = covariates,
  target = "USD_unitcost_total",
  priors = priors
)
```

Priors can be visualised by calling `plot(priors, parameter = "name_of_param")`:

```{r beta-priors, fig.width=10, fig.height=6}
library(gridExtra)
priors <- model$priors()
plots <- list()
for(i in 1:3) {
  plots[[i]] <- plot(priors, parameter = paste0("beta[", i, "]"))
}
plots[[4]] <- plot(priors, parameter = "mu_alpha")
plots[[5]] <- plot(priors, parameter = "sigma_alpha")
plots[[6]] <- plot(priors, parameter = "sigma")
do.call(grid.arrange, c(plots, ncol = 3))
```

## Fitting the model

Under the hood the MixedEffects class uses JAGS (Just Another Gibbs Sampler) to fit a hierarchical linear regression model with country-specific intercepts and shared covariate effects. To fit the model, [JAGS](https://sourceforge.net/projects/mcmc-jags/) and the [rjags](https://cran.r-project.org/web/packages/rjags/index.html) package must be installed on your machine.

For the purpose of the vignette, we'll use fewer iterations than recommended for faster computation:

```{r model-fitting, results='hide', message=FALSE}
# Fit the model with reduced iterations for demonstration
# In practice, you might want to use the defaults (n.iter = 100000)
model$fit(
 n.iter = 20000,
 n.burnin = 1000,
 n.thin = 5,
 n.chains = 2
)
```

```{r model-summary}
# Check summary statistics of the fitted samples
fitted_samples <- model$samples()
summary(fitted_samples)
```

### Diagnostics

Several functions are available on the model class to check whether the model has converged:

Rhat:

```{r}
model$mcmc_rhat()
```

Trace:

```{r}
model$mcmc_trace()
```

Auto-correlation:

```{r}
model$mcmc_acf()
```

Effective sample size:

```{r}
knitr::kable(model$n_eff())
```

We can also plot the posterior distributions of each parameter:

```{r}
model$plot_posteriors(pars = paste0("beta[", 1:length(covariates), "]")) + 
  scale_y_discrete(labels = covariates)
```

```{r}
model$plot_posteriors(pars = c("mu_alpha", "sigma_alpha", "sigma",  paste0("alpha[", 1:5, "]")))
```

## Generating predictions

We can now generate predictions for the training data and evaluate model fit:

```{r predictions}
# Generate predictions for the training data on log scale
dat <- model$training_data()
predictions <- model$predict(dat, scale = "log", summarised = TRUE)
head(predictions)

# Various measures of fit
performance <- model$performance(scale = "log")
knitr::kable(performance)
```

## Visualising results

### 1. Predicted vs Observed

```{r scatter-plot}
# Create scatter plot of predicted vs observed values
model$plot_fit(include_ci = FALSE, scale = "log")
```

### 2. 95% Credible Prediction Intervals

```{r prediction-intervals}
# Plot with prediction intervals
model$plot_fit(include_ci = TRUE, scale = "log")
```

### 3. Residuals

```{r residuals}
model$plot_residuals(add_smooth = TRUE, color_by_country = TRUE)
```

### 4. Country-Specific perfomance

```{r country-analysis}
# Performance by country
country_performance <- model$performance(by_country = TRUE)
colnames(country_performance) <- c("Country", "MAE",
 "RMSE", "Correlation", "95% CI Coverage")

knitr::kable(country_performance)

knitr::kable(country_performance)
```

```{r country-plot}
model$plot_fit() + 
  facet_wrap(~country, scales = "free")
```

### 5. Out-of-sample performance

We can check for overfitting and estimate out-of-sample performance using k-fold cross-validation. Here we use 3 folds for quick compilation; in practice, 10 or 20 folds would give a more accurate picture.

```{r}
res <- model$k_fold_cv(k_folds = 3, 
  n.iter = 10000,
  n.burnin = 1000,
  n.adapt = 1000, 
  scale = "log")

fit <- res |> 
  group_by(fold) |> 
  summarise(rmse = sqrt(mean((observed-mean)^2)),
            mae = mean(abs(observed-mean)))

knitr::kable(fit)
```
