---
title: "Model Fitting"
author: "Alex Hill"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Model Fitting}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

```{r setup}
library(capturetb)
library(ggplot2)
library(dplyr)
```

## Introduction

This vignette demonstrates how to use this package to:

1. Create a new mixed effects model instance with covariates, priors and training data
2. Fit the model to estimate parameters

This demonstrates how the default model was fitted and sensitivity analyses performed.

Under the hood this uses JAGS (Just Another Gibbs Sampler) to fit a hierarchical linear regression model with country-specific intercepts and shared covariate effects.

## Creating a model instance

A model instance requires a list of covariates, a target variable to predict, training data, and priors for parameters. 

```{r model-creation}
covariates <- c("log_USD_p_bldgspace",
 "logVisits", "logVisitsPP_TB", "primary",
 "secondary", "urban", "public")

# Specifying priors for the fixed effects
# Others will take default values
priors <- capturetb::capturetb_priors(
  beta.mean = c(0,-0.0142, -0.0412, 0.304, 0.348, 0.352, -0.29),
  beta.precision = c(0.01, 227, 139, 46.4, 80.8, 55.2, 234.5))

data <- get_data("OP diagnostic visit") # or provide your own data; see capturetb::outputs() for all output data installed with the package

model <- capturetb::MixedEffects$new(
  dat = data,
  covariates = covariates,
  target = "USD_unitcost_total",
  priors = priors
)

# See all priors used in the model
priors <- model$priors()
print(priors)
```

Priors can be visualised by calling `plot(priors, parameter = "name_of_param")`:

```{r beta-priors, fig.width=10, fig.height=6}
library(gridExtra)
plots <- list()
for(i in 1:3) {
  plots[[i]] <- plot(priors, parameter = paste0("beta[", i, "]"))
}
plots[[4]] <- plot(priors, parameter = "mu_alpha_mean")
plots[[5]] <- plot(priors, parameter = "sigma_alpha")
plots[[6]] <- plot(priors, parameter = "sigma")
do.call(grid.arrange, c(plots, ncol = 3))
```

## Fitting the model

For the purpose of the vignette, we'll use fewer iterations than recommended for faster computation:

```{r model-fitting, results='hide', message=FALSE}
# Fit the model with reduced iterations for demonstration
# In practice, you might want to use the defaults (n.iter = 100000)
model$fit(
 n.iter = 20000,
 n.burnin = 1000,
 n.thin = 5,
 n.chains = 2
)
```

```{r model-summary}
# Check summary statistics of the fitted samples
fitted_samples <- model$samples()
summary(fitted_samples)
```

### Diagnostics

Rhat:

```{r}
model$mcmc_rhat()
```

Trace:

```{r}
model$mcmc_trace()
```

Auto-correlation:

```{r}
model$mcmc_acf()
```

Effective sample size:

```{r}
knitr::kable(model$n_eff())
```

Plotting posteriors:

```{r}
model$plot_posteriors(pars = paste0("beta[", 1:length(covariates), "]")) + 
  scale_y_discrete(labels = covariates)
```

```{r}
model$plot_posteriors(pars = c("mu_alpha", "sigma_alpha", "sigma",  paste0("alpha[", 1:5, "]")))
```

## Generating predictions

We can now generate predictions for the training data to evaluate model fit:

```{r predictions}
# Generate predictions for the training data on log scale
dat <- model$training_data()
predictions <- model$predict(dat, scale = "log", summarised = TRUE)
head(predictions)

# Various measures of fit
performance <- model$performance(scale = "log")
knitr::kable(performance)
```

## Visualising results

### 1. Predicted vs Observed

```{r scatter-plot}
# Create scatter plot of predicted vs observed values
model$plot_fit(include_ci = FALSE, scale = "log")
```

### 2. 95% Credible Prediction Intervals

```{r prediction-intervals}
# Plot with prediction intervals
model$plot_fit(include_ci = "TRUE", scale = "log")
```

### 3. Residuals

```{r residuals}
model$plot_residuals(add_smooth = TRUE, color_by_country = TRUE)
```

### 4. Country-Specific perfomance

```{r country-analysis}
# Performance by country
results_df <- cbind(
  observed = log(dat$USD_unitcost_total),
  country = dat$fc_country,
  predictions
)

country_performance <- results_df %>%
  group_by(country) %>%
  summarise(
    n_obs = n(),
    mae = mean(abs(observed - mean)),
    rmse = sqrt(mean((observed - mean)^2)),
    correlation = cor(observed, mean),
    coverage = mean(observed >= lower & observed <= upper),
    .groups = 'drop'
  ) %>%
  arrange(desc(n_obs))
colnames(country_performance) <- c("Country", "N", "MAE",
 "RMSE", "Corelation", "95% CI Coverage")

knitr::kable(country_performance)
```

```{r country-plot}
model$plot_fit() + 
  facet_wrap(~country, scales = "free")
```

### 5. Out-of-sample performance

We can check for overfitting and estimate out-of-sample performance using k-fold cross-validation:

```{r}
res <- model$k_fold_cv(k_folds = 3, 
  n.iter = 10000,
  n.burnin = 1000,
  n.adapt = 1000, 
  scale = "log")

fit <- res |> 
  group_by(fold) |> 
  summarise(rmse = sqrt(mean((observed-mean)^2)),
            mae = mean(abs(observed-mean)))

knitr::kable(fit)
```
