---
title: "Model Fitting and Prediction with CaptureTB"
author: "Alex Hill"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Model Fitting and Prediction with CaptureTB}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

```{r setup}
library(capturetb)
library(ggplot2)
library(dplyr)
```

## Introduction

This vignette demonstrates how to use the CaptureTB package to:

1. Fit a Bayesian hierarchical model to tuberculosis cost data
2. Generate predictions from the fitted model
3. Compare predicted values to observed values

The package uses JAGS (Just Another Gibbs Sampler) to fit a hierarchical linear regression model with country-specific intercepts and shared covariate effects.

## Loading and Exploring the Data

First, let's load the data and explore what's available:

```{r data-exploration}
# Get all available output types
available_outputs <- outputs()
print(available_outputs)

# Load data for outpatient treatment visits
data <- get_data("OP treatment visit")
print(paste("Number of observations:", nrow(data)))
print(paste("Number of countries:", length(unique(data$fc_country))))

# View the structure of the data
str(data[, c("fc_country", "USD_unitcost_total", capturetb_covariates())])
```

## Setting Up Priors

By default, the model uses power priors derived from the WHO-CHOICE model of outpatient treatment visit costs.

```{r priors}
# Create default priors
priors <- capturetb_priors()
knitr::kable(as.data.frame(priors))
```

Priors can be plotted by calling `plot(priors, parameter = "name_of_param")`:

```{r beta-priors, fig.width=10, fig.height=6}
# Plot prior distributions for the first few beta coefficients
library(gridExtra)
plots <- list()
for(i in 1:3) {
  plots[[i]] <- plot(priors, parameter = paste0("beta[", i, "]"))
}
do.call(grid.arrange, c(plots, ncol = 3))
```

## Fitting the Model

Now we can fit the Bayesian hierarchical model. For the purpose of the vignette, we'll use fewer iterations that recommended for faster computation:

```{r model-fitting, results='hide', message=FALSE}
# Fit the model with reduced iterations for demonstration
# In practice, you might want to use the defaults (n.iter = 100000)
fitted_model <- fit_capturetb_model(
  dat = data,
  covariates = capturetb_covariates(),
  target = "USD_unitcost_total",
  priors = priors,
  n.iter = 10000,
  n.burnin = 500,
  n.thin = 5,
  seed = 123
)
```

```{r model-summary}
# Check summary statistics
summary(fitted_model)
```

Checking Rhat values to see if the model coverged:

```{r}
gr <- coda::gelman.diag(fitted_model,
                  autoburnin = FALSE,
                  multivariate = FALSE)
          
ggplot(data.frame(Parameter = names(gr$psrf[,1]),
                  Rhat      = gr$psrf[, "Point est."]),
       aes(x = reorder(Parameter, Rhat), y = Rhat)) +
  geom_hline(yintercept = 1.01, linetype = "dashed") +
  geom_point(col = "red") +
  labs(y = expression(hat(R)), x = NULL) +
  theme_minimal()
```

## Generating Predictions

We can now generate predictions for the training data to evaluate model fit:

```{r predictions}
# Get the covariates and countries from the training data
covariates <- capturetb_covariates()
countries <- unique(data$fc_country)

# Generate predictions for the training data
log_predictions <- predict_capturetb_logcost(
  dat = data,
  covariates = covariates,
  countries = countries,
  samples = fitted_model
)

# Convert to cost scale (exponentiate)
predictions <- exp(log_predictions)
predictions <- predictions[rowSums(is.infinite(predictions) == 0), , drop = FALSE ]

# Calculate summary statistics for predictions
pred_mean <- apply(predictions, 2, mean)
pred_lower <- apply(predictions, 2, quantile, 0.025)
pred_upper <- apply(predictions, 2, quantile, 0.975)

# Create a data frame with predictions and observed values
results_df <- data.frame(
  observed = data$USD_unitcost_total,
  predicted_mean = pred_mean,
  predicted_lower = pred_lower,
  predicted_upper = pred_upper,
  country = data$fc_country
)
```

## Plotting Predicted vs Observed Values

### 1. Predicted vs Observed Scatter Plot

```{r scatter-plot}
# Create scatter plot of predicted vs observed values
ggplot(results_df, aes(x = observed, y = predicted_mean)) +
  geom_point(aes(color = country), alpha = 0.7) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "blue") +
  scale_x_log10() +
  scale_y_log10() +
  labs(
    title = "Predicted vs Observed Unit Costs",
    x = "Observed Cost (USD)",
    y = "Predicted Cost (USD)",
    color = "Country"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")
```

### 2. 95% Credible Prediction Intervals

```{r prediction-intervals}
# Plot with prediction intervals
ggplot(results_df, aes(x = observed)) +
  geom_errorbar(aes(ymin = predicted_lower, ymax = predicted_upper), 
              alpha = 0.3) +
  geom_point(aes(y = predicted_mean, colour = country), alpha = 0.7) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "blue") +
  scale_y_log10() +
  scale_x_log10() +
  labs(
    title = "Credible Intervals vs Observed Unit Costs",
    x = "Observed Cost (USD)",
    y = "Predicted Cost (USD)",
    color = "Country"
  ) +
  theme_minimal()
```

### 3. Residuals

```{r residuals}
# Calculate residuals on log scale
results_df$residuals <- log(results_df$observed) - log(results_df$predicted_mean)

# Plot residuals
ggplot(results_df, aes(x = log(predicted_mean), y = residuals)) +
  geom_point(aes(color = country), alpha = 0.7) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  geom_smooth(method = "loess", se = TRUE, color = "grey", alpha = 0.3) +
  labs(
    title = "Residuals vs Fitted Values",
    x = "Log Predicted Cost",
    y = "Residuals (Log Observed - Log Predicted)",
    color = "Country"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")
```

### 4. Model Performance Metrics

```{r performance-metrics}
performance <- results_df %>%
  summarise(
    mae = mean(abs(observed - predicted_mean)),
    rmse = sqrt(mean((observed - predicted_mean)^2)),
    correlation = cor(observed, predicted_mean),
    coverage = mean(observed >= predicted_lower & observed <= predicted_upper)
  ) 
colnames(performance) <- c("MAE", "RMSE", "Corelation", "95% CI Coverage")
knitr::kable(performance)
```

## Country-Specific Perfomance

Let's examine how well the model performs for different countries:

```{r country-analysis}
# Performance by country
country_performance <- results_df %>%
  group_by(country) %>%
  summarise(
    n_obs = n(),
    mae = mean(abs(observed - predicted_mean)),
    rmse = sqrt(mean((observed - predicted_mean)^2)),
    correlation = cor(observed, predicted_mean),
    coverage = mean(observed >= predicted_lower & observed <= predicted_upper),
    .groups = 'drop'
  ) %>%
  arrange(desc(n_obs))
colnames(country_performance) <- c("Country", "N", "MAE", "RMSE", "Corelation", "95% CI Coverage")

knitr::kable(country_performance)
```

```{r country-plot}
ggplot(results_df, aes(x = observed, y = predicted_mean)) +
  geom_point(alpha = 0.7) +
  geom_errorbar(aes(ymin = predicted_lower, ymax = predicted_upper), 
              alpha = 0.3) +
  scale_y_log10() +
  scale_x_log10() +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "blue") +
  facet_wrap(~country, scales = "free") +
  labs(
    title = "Predicted vs Observed by Country",
    x = "Observed Cost (USD, log scale)",
    y = "Predicted Cost (USD, log scale)"
  ) +
  theme_minimal()
```
