---
title: "Model Comparisons"
author: "Alex Hill"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Model Comparisons}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

```{r setup}
library(capturetb)
library(ggplot2)
library(dplyr)
```

## Model Comparisons

This vignette compares 3 model structures. In each structure, log costs for a particular set of input variables are assumed to follow a normal distribution:

$$
\log(\text{USD\_unitcost\_total}_i) \sim \mathcal{N}(\mu_i, \sigma^2)
$$

where $\mu_i$ is a linear combination of the inputs.

1. Fixed effects - all parameters assumed the same across observations

$$
\mu_i = \alpha + \sum_{j=1}^J{\beta_j \times x_{ji}}
$$

2. Mixed effects - intercepts vary by country
$$
\mu_i = \alpha_{c_i} + \sum_{j=1}^J{\beta_j \times x_{ji}}
$$

where 

$$
\alpha_{c} \sim \mathcal{N}(\mu_\alpha, \sigma_\alpha^2)
$$

3. Random slopes - intercepts and slopes vary by country

$$
\mu_i = \alpha_{c_i} + \sum_{j=1}^J{\beta_{jc_i} \times x_{ji}}
$$

where also, for each j,

$$
\beta_j \sim \mathcal{N}(\mu_{\beta_j}, \sigma_{\beta_j}^2)
$$

We compare the estimated effects of each covariate when fit using the different model structures:

```{r}
library(coda)
# get samples from pre-fitted model
mixed_effects <- capturetb::unitcost()
samples_mixed <- as.matrix(mixed_effects$samples())

# fit fixed effects model to same data and priors
fixed_effects <- capturetb::FixedEffects$new()
fixed_effects$fit(quiet = TRUE)
samples_fixed <- as.matrix(fixed_effects$samples())

# fit random effects model to same data and priors
random_effects <- capturetb::RandomSlopes$new()
random_effects$fit(quiet = TRUE)
samples_random <- as.matrix(random_effects$samples())
```

```{r, samples, echo = FALSE}
#| fig-cap: Comparison of intercept estimates across models
#| fig-width: 6
#| fig-height: 6
#| fig-align: center
countries <- random_effects$countries()
covariates <- random_effects$covariates()

lower <- function(x) quantile(x, probs = 0.025)
upper <- function(x) quantile(x, probs = 0.975)

alpha_cols <- paste0("alpha[", seq_along(countries), "]")
beta_cols <- with(expand.grid(i = seq_along(covariates), j = seq_along(countries)),
               sprintf("beta[%d,%d]", i, j))
mu_beta_cols <- paste0("mu_beta[", seq_along(covariates), "]")

alpha_post_random <- as.matrix(samples_random[, alpha_cols, drop = FALSE])
beta_post_random <- as.matrix(samples_random[, beta_cols, drop = FALSE])
mu_beta_post_random <- as.matrix(samples_random[, mu_beta_cols, drop = FALSE])

beta_cols <- paste0("beta[", seq_along(covariates), "]")
alpha_post_mixed <- as.matrix(samples_mixed[, alpha_cols, drop = FALSE])
beta_post_mixed <- as.matrix(samples_mixed[, beta_cols, drop = FALSE])

beta_post_fixed <- as.matrix(samples_fixed[, beta_cols, drop = FALSE])
alpha_post_fixed <- as.matrix(samples_fixed[, "alpha", drop = FALSE])

# betas
beta_mean_fixed <- apply(beta_post_fixed, 2, mean)
beta_lo_fixed <- apply(beta_post_fixed, 2, lower)
beta_hi_fixed <- apply(beta_post_fixed, 2, upper)

beta_mean_mixed <- apply(beta_post_mixed, 2, mean)
beta_lo_mixed <- apply(beta_post_mixed, 2, lower)
beta_hi_mixed <- apply(beta_post_mixed, 2, upper)

beta_mean_random <- matrix(apply(beta_post_random, 2, mean), ncol = 6, byrow = TRUE)
beta_lo_random <- matrix(apply(beta_post_random, 2, lower), ncol= 6, byrow = TRUE)
beta_hi_random <- matrix(apply(beta_post_random, 2, upper), ncol = 6, byrow = TRUE)
mu_beta_mean <- apply(mu_beta_post_random, 2, mean)
mu_beta_lo <- apply(mu_beta_post_random, 2, lower)
mu_beta_hi <- apply(mu_beta_post_random, 2, upper)

# alphas 

alpha_lo <- apply(alpha_post_fixed, 2, lower)
alpha_hi <- apply(alpha_post_fixed, 2, upper)
alpha_mean <- apply(alpha_post_fixed, 2, mean)

alpha_mean_random <- apply(alpha_post_random, 2, mean)
alpha_lo_random <- apply(alpha_post_random, 2, lower)
alpha_hi_random <- apply(alpha_post_random, 2, upper)

alpha_mean_mixed <- apply(alpha_post_mixed, 2, mean)
alpha_lo_mixed <- apply(alpha_post_mixed, 2, lower)
alpha_hi_mixed <- apply(alpha_post_mixed, 2, upper)

library(ggplot2)

ggplot(data.frame(
	cat = rep(countries, 3),
	Model = c(
		rep("Fixed effects", 5),
	 	rep("Country-specific intercepts", 5), 
		rep("Country-specific slopes", 5)),
	lo = c(rep(alpha_lo, 5), alpha_lo_mixed, alpha_lo_random),
	hi = c(rep(alpha_hi, 5), alpha_hi_mixed, alpha_hi_random),
 	val = c(rep(alpha_mean, 5), alpha_mean_mixed, alpha_mean_random))) + 
	geom_point(position = position_dodge(width = 0.5), 
		aes(x = cat, y = val, col = Model)) +
	geom_errorbar(aes(x = cat,
	 	ymin = lo,
		ymax = hi, 
		col = Model),
		position = position_dodge(width = 0.5)) +
	labs(y = NULL, x = NULL) + 
	theme(axis.text.x = element_text(angle = 45, hjust = 1), 
		legend.position = "bottom")
```

Comparing fixed coefficient estimates between the first 2 models and population mean coefficients from the third:

```{r, coeffs-plot-compare, echo = FALSE}
#| fig-width: 6
#| fig-height: 6
#| fig-align: center
#| fig-cap: Comparison of coefficients across models
 
vars <- covariates
ggplot(data.frame(
	cat = rep(vars, 3),
	Model = c(
		rep("Fixed effects", 6), 
		rep("Country-specific intercepts", 6),
	 	rep("Country-specific slopes", 6)),
	lo = c(beta_lo_fixed, beta_lo_mixed, mu_beta_lo),
	hi = c(beta_hi_fixed, beta_hi_mixed, mu_beta_hi),
 	val = c(beta_mean_fixed, beta_mean_mixed, mu_beta_mean))) + 
	geom_point(position = position_dodge(width = 0.5), 
		aes(x = cat, y = val, col = Model)) +
	geom_errorbar(aes(x = cat, ymin = lo, ymax = hi, col = Model),
		position = position_dodge(width = 0.5)) +
	labs(y = NULL, x = NULL) + 
	theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "bottom")
```

Comparing coefficient estimates between the fixed effects model and country-specific coefficients model:

```{r, coeffs1, echo = FALSE}
#| fig-width: 10
#| fig-height: 10
#| fig-align: center
#| fig-cap: Comparison of country-specific slopes and fixed effects model coefficients, by country
 
drawCoeffPlot <- function(i) {
	variable <- covariates[[i]]

	ggplot() + geom_point(aes(y = countries, x = beta_mean_random[, i])) +
		geom_vline(xintercept = beta_mean_fixed[i], linetype = "dashed") +
		geom_rect(aes(xmin = beta_lo_fixed[i], 
			xmax = beta_hi_fixed[i], 
			ymin = -Inf, 
			ymax = Inf), 
			alpha = 0.2) +
		geom_segment(aes(x = beta_lo_random[, i], 
			xend = beta_hi_random[, i], 
			y = countries, 
			yend = countries), 
			alpha = 0.8) + 
		labs(title = variable, x = NULL, y = NULL)
}

gridExtra::grid.arrange(grobs = lapply(1:6, drawCoeffPlot))
```

And between the mixed effects model and country-specific slopes model:

```{r, coeffs2, echo = FALSE}
#| fig-width: 10
#| fig-height: 10
#| fig-align: center
#| fig-cap: Comparison of country-specific and mixed effects model coefficients, by country
 
drawCoeffPlot <- function(i) {
	variable <- covariates[[i]]

	ggplot() + geom_point(aes(y = countries, x = beta_mean_random[, i])) +
		geom_vline(xintercept = beta_mean_mixed[i], linetype = "dashed") +
		geom_rect(aes(xmin = beta_lo_mixed[i], 
			xmax = beta_hi_mixed[i], 
			ymin = -Inf, 
			ymax = Inf), 
			alpha = 0.2) +
		geom_segment(aes(x = beta_lo_random[, i], 
			xend = beta_hi_random[, i], 
			y = countries, 
			yend = countries), 
			alpha = 0.8) + 
		labs(title = variable, x = NULL, y = NULL)
}

gridExtra::grid.arrange(grobs = lapply(1:6, drawCoeffPlot))
```

Comparing metrics of fit for all 3:

```{r}
knitr::kable(rbind(
    fixed = fixed_effects$performance(),
    mixed = mixed_effects$performance(),
    random = random_effects$performance()
))
```

Visualising predictions:

```{r}
#| fig-width: 10
#| fig-height: 10

gridExtra::grid.arrange(grobs = list(
	fixed_effects$plot_fit(),
    mixed_effects$plot_fit(),
    random_effects$plot_fit()))
```

Checking each for over-fitting:

```{r}
library(dplyr)
kfold_fixed <- fixed_effects$k_fold_cv(k_folds = 20, 
  n.iter = 50000,
  n.burnin = 1000,
  n.adapt = 1000, 
  seed = 123,
  quiet = TRUE,
  scale = "log")

fit_fixed <- kfold_fixed |> 
  group_by(fold) |> 
  summarise(rmse = sqrt(mean((observed-mean)^2)))

rmse_fixed <- sqrt(mean((kfold_fixed$observed - kfold_fixed$mean)^2))

kfold_mixed <- mixed_effects$k_fold_cv(k_folds = 20, 
  n.iter = 50000,
  n.burnin = 1000,
  n.adapt = 1000, 
  seed = 123,
  quiet = TRUE,
  scale = "log")

fit_mixed <- kfold_mixed |> 
  group_by(fold) |> 
  summarise(rmse = sqrt(mean((observed-mean)^2)))

rmse_mixed <- sqrt(mean((kfold_mixed$observed - kfold_mixed$mean)^2))

kfold_random <- random_effects$k_fold_cv(k_folds = 20, 
  n.iter = 50000,
  n.burnin = 1000,
  n.adapt = 1000,
  seed = 123, 
  quiet = TRUE,
  scale = "log")

rmse_random <- sqrt(mean((kfold_random$observed - kfold_random$mean)^2))

fit_random <- kfold_random |> 
  group_by(fold) |> 
  summarise(rmse = sqrt(mean((observed-mean)^2)))

# Out of sample RMSE
knitr::kable(cbind(fixed = rmse_fixed, mixed = rmse_mixed, random = rmse_random))
```

```{r}
fits <- rbind(
    cbind(mod = "fixed", fit_fixed),
    cbind(mod = "mixed", fit_mixed), 
    cbind(mod = "random", fit_random)
)
ggplot(fits, aes(x = mod, y = rmse)) + geom_boxplot()
```

As with in-sample fit, the mixed effects and random slopes model show significantly better performance than the fixed effects model. Between the mixed effects and random slopes model, there's not much in it, although there does seem to be less variance in the error of the mixed effects model across folds. 

Comparing country-specific performance of the models:

```{r}
fe_country <- cbind(mod = "fixed", fixed_effects$performance(by_country = TRUE))
me_country <- cbind(mod = "mixed", mixed_effects$performance(by_country = TRUE))
re_country <- cbind(mod = "random", random_effects$performance(by_country = TRUE))
knitr::kable(rbind(fe_country, me_country, re_country))
```
