---
title: "Using the model"
author: "Alex Hill"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{using the model}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

```{r setup}
library(capturetb)
library(ggplot2)
library(dplyr)
```

## Introduction

This vignette demonstrates how to use this package to:

1. Generate unit cost predictions from the fitted model
2. Compare predicted values to observed values to assess model fit

## Loading and exploring the raw data

Top-down economic costs from the ValueTB database are installed with the package. This includes costs for multiple different output types; run `capturetb::outputs()` to retrieve a complete list of available outputs:

```{r data-exploration}
available_outputs <- capturetb::outputs()
head(available_outputs)
```

Subsets of the data can be retrieved using `capturetb::get_data()`.

```{r load-data}
data <- capturetb::get_data("OP treatment visit")
summarised <- data |> 
  group_by(fc_country) |>
  summarise(n_obs = n())

knitr::kable(summarised)
```

This package includes fitted models for predicting the following outputs:
* OP treatment visit unit cost
* OP treatment visit fixed unit cost (average fixed costs per visit)

## Exploring the model

 The OP treatment visit unit costs model is loaded via `capturetb_total`:

```{r model-creation}
model <- capturetb::unitcost()

# See model covariates
covariates <- model$covariates()
print(covariates)

# See priors used in the model
priors <- model$priors()
print(priors)
```

Priors can be visualised by calling `plot(priors, parameter = "name_of_param")`:

```{r beta-priors, fig.width=10, fig.height=6}
library(gridExtra)
plots <- list()
for(i in 1:3) {
  plots[[i]] <- plot(priors, parameter = paste0("beta[", i, "]"))
}
plots[[4]] <- plot(priors, parameter = "mu_alpha_mean")
plots[[5]] <- plot(priors, parameter = "sigma_alpha")
plots[[6]] <- plot(priors, parameter = "sigma")
do.call(grid.arrange, c(plots, ncol = 3))
```

Plotting posteriors:

```{r}
fitted_samples <- model$samples()
bayesplot::mcmc_areas(fitted_samples, 
                      pars = paste0("beta[", 1:length(covariates), "]"),
                      prob = 0.9) + 
  scale_y_discrete(labels = covariates)
```

```{r}
bayesplot::mcmc_areas(fitted_samples, 
                      pars = c("mu_alpha", "sigma_alpha", "sigma",  paste0("alpha[", 1:5, "]")),
                      prob = 0.9)
```

## Generating predictions

We can generate predictions for the training data to demonstrate model fit:

```{r predictions}
# Generate predictions for the training data on log scale
dat <- model$training_data()
predictions <- model$predict(dat, scale = "log")

# Calculate summary statistics for predictions
pred_mean <- apply(predictions, 2, mean)
pred_lower <- apply(predictions, 2, quantile, 0.025)
pred_upper <- apply(predictions, 2, quantile, 0.975)

# Create a data frame with predictions and observed values
results_df <- data.frame(
  observed = log(dat$USD_unitcost_total),
  predicted_mean = pred_mean,
  predicted_lower = pred_lower,
  predicted_upper = pred_upper,
  country = dat$fc_country
)

# Various measures of fit
performance <- results_df %>%
  summarise(
    mae = mean(abs(observed - predicted_mean)),
    rmse = sqrt(mean((observed - predicted_mean)^2)),
    correlation = cor(observed, predicted_mean),
    coverage = mean(observed >= predicted_lower & observed <= predicted_upper)
  ) 
colnames(performance) <- c("MAE", "RMSE", "Correlation", "95% CI Coverage")
knitr::kable(performance)
```

Predictions of unit cost for new inputs, not in the training set, are generated in exactly the same way, via the `predict` method.

## Visualising results

### 1. Predicted vs observed

```{r scatter-plot}
# Create scatter plot of predicted vs observed values
ggplot(results_df, aes(x = observed, y = predicted_mean)) +
  geom_point(aes(color = country), alpha = 0.7) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "blue") +
  labs(
    title = "Predicted vs Observed Unit Costs",
    x = "Observed Cost (log(USD))",
    y = "Predicted Cost (log(USD))",
    color = "Country"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")
```

### 2. 95% Credible Prediction Intervals

```{r prediction-intervals}
# Plot with prediction intervals
ggplot(results_df, aes(x = observed)) +
  geom_errorbar(aes(ymin = predicted_lower, ymax = predicted_upper), 
              alpha = 0.3) +
  geom_point(aes(y = predicted_mean, colour = country), alpha = 0.7) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "blue") +
  labs(
    title = "Credible Intervals vs Observed Unit Costs",
    x = "Observed Cost (log(USD))",
    y = "Predicted Cost (log(USD))",
    color = "Country"
  ) +
  theme_minimal()
```

### 3. Country-Specific perfomance

```{r country-analysis}
# Performance by country
country_performance <- results_df %>%
  group_by(country) %>%
  summarise(
    n_obs = n(),
    mae = mean(abs(observed - predicted_mean)),
    rmse = sqrt(mean((observed - predicted_mean)^2)),
    correlation = cor(observed, predicted_mean),
    coverage = mean(observed >= predicted_lower & observed <= predicted_upper),
    .groups = 'drop'
  ) %>%
  arrange(desc(n_obs))
colnames(country_performance) <- c("Country", "N", "MAE",
 "RMSE", "Corelation", "95% CI Coverage")

knitr::kable(country_performance)
```

```{r country-plot}
ggplot(results_df, aes(x = observed, y = predicted_mean)) +
  geom_point(alpha = 0.7) +
  geom_errorbar(aes(ymin = predicted_lower, ymax = predicted_upper), 
              alpha = 0.3) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "blue") +
  facet_wrap(~country, scales = "free") +
  labs(
    title = "Predicted vs Observed by Country",
    x = "Observed Cost (USD, log scale)",
    y = "Predicted Cost (USD, log scale)"
  ) +
  theme_minimal()
```
